\begin{center}
    {\LARGE \bf CURRICULUM VITAE}\\[0.5in]
    {\large \bf Hasung Song}\\[0.1in]
    Department of Physics\\
    Boston University\\
    Email: hwsong@bu.edu\\
    % Add more contact info as needed
\end{center}

\vspace{0.3in}

\section*{Education}
\begin{itemize}[leftmargin=*]
    \item \textbf{Ph.D. in Physics}, Boston University, 2019--Present
    \item \textbf{M.Sc. in Physics}, Boston University, 2019--2022
    \item \textbf{B.Sc. in Physics}, University of California, Berkeley, 2016-2018
\end{itemize}

\section*{Professional Experience}
\begin{itemize}[leftmargin=*]
    \item \textbf{Post-Baccalaureate Researcher}, University of California, Berkeley, 2018--2019
\end{itemize}

\section*{Selected Publications}
\begin{itemize}[leftmargin=*]
	\item "Search for Majorana Neutrinos with the Complete KamLAND Zen Dataset" KamLAND-Zen Collaboration,	arXiv:2406.11438 [hep-ex] (2024)
	\item "Eos: conceptual design for a demonstrator of hybrid optical detector technology" T. Anderson et al. JINST 18 P02009 (2023)
	\item "KamNet: An integrated spatiotemporal deep neural network for rare event searches in KamLAND-Zen" A. Li, et al. Phys. Rev. C 107, 014323 (2023)
	\item "RFSoC-based front-end electronics for pulse detection" S.N. Axani et al. 2024 JINST 19 P03013 (2024)
	\item "Search for the Majorana Nature of Neutrinos in the Inverted Mass Ordering Region with KamLAND-Zen" KamLAND-Zen Collaboration, Phys. Rev. Lett. 130, 051801 (2023)
\end{itemize}

\section*{Scientific Collaborations}
\begin{itemize}[leftmargin=*]
	\item EoS Collaboration 2023--Present
	\item NuDOT Collaboration 2022--Present
    \item KamLAND Collaboration 2020--Present
    \item NIM+ 2020--2021
	\item Mu2E Collaboration 2018--2019
\end{itemize}

\section*{Conferences and Presentations}
\begin{itemize}[leftmargin=*]
    \item "Getting Excited About Double Beta Decay in KamLAND-Zen", APS DNP, Talk, Boston, MA, 2024.
    \item "Understanding KamNet Performance in KamLAND-ZEN with ML Interpretability", SLAC Summer Institute, Poster, SLAC, 2023.
    \item "KamLAND-Zen's New ML Tricks", Neutrino Physics and Machine Learning, Talk, Northeastern University, 2023.
    \item "Rejecting Spallation Backgrounds in KamLAND-ZEN with KamNet", Neutrino, Poster, Seoul, Korea, 2022.
    \item "Rejecting Spallation Backgrounds in KamLAND-ZEN with KamNet", TAUP, Poster, Valenica, Spain, 2021.
    \item "Neutrinoless Muon-to-Positron Conversion at Mu2e", APS April Meeting, Talk, Denver, Colorado, 2019.
\end{itemize}

\section*{Teaching Experience}
\begin{itemize}[leftmargin=*]
    \item Physics 211 General Physics I, Teaching Assistant, Boston University, Summer 2020.
    \item Physics 211 General Physics I, Teaching Assistant, Boston University, Fall 2019.
\end{itemize}

\section*{Professional Service}
\begin{itemize}[leftmargin=*]
    \item Member, Graduate Student Council, Boston University Physics Department, 2019--Present.
\end{itemize}

\end{document}
